---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%",
  dev = "svglite"
)
```

# dltam

<!-- badges: start -->
<!-- badges: end -->

This package implements the calculation of the dynamic local trend association measure (DLTAM) for
a pair of time series, proposed by García-López, F. J., Batyrshin, I., and Gelbukh, A., in the paper
["Dynamic Local Trend Associations in Analysis of Comovements of Financial Time Series"](https://doi.org/10.1007/978-3-319-67137-6_20).

## Installation

You can install this package using:

``` r
# install.packages("remotes")
remotes::install_github("adamoshen/dltam")
```

## Basic usage

```{r}
library(tidyverse)
library(patchwork)
library(loco)
library(dltam)
```

Suppose we are interested in the DLTAM of average daily temperature readings at JFK and LAX
airports. This data can be found in the [loco](https://github.com/adamoshen/loco) package. A quick
preview of the data:

```{r}
daily_temperature %>%
  group_by(origin) %>%
  slice_head(n = 6) %>%
  group_split()
```

First, we require a data frame containing the two time series' data as columns.

```{r}
daily_temperature_wide <- daily_temperature %>%
  pivot_wider(id_cols=date, names_from=origin, values_from=temp)

daily_temperature_wide
```

The DLTAM scores can be obtained by passing `daily_temperature_wide` to `dltam::dltam()`. Following
the example in the paper, which used daily data, we set the window size to 30 and the Moving
Approximation Transform sequence size to 16.

```{r}
dltam_scores <- daily_temperature_wide %>%
  dltam(JFK, LAX, timestamp=date, window_size=30, mat_seq_size=16)

dltam_scores
```

We can now plot the DLTAM scores and identify time points of particular interest.

```{r, fig.alt="Plot 1: Daily average temperatures in fahrenheit in 2013 for JFK and LAX airports. Plot 2: DLTAM scores."}
p1 <- ggplot(daily_temperature) +
  geom_line(aes(x=date, y=temp, colour=origin), alpha=0.6) +
  scale_colour_manual(values = c("#009E73", "#D55E00")) +
  scale_x_date(date_breaks="2 months", date_labels="%b") +
  labs(x="", y="Average daily temperature (°F)", colour="") +
  theme_bw(base_size = 9) +
  theme(legend.position = "bottom")

p2 <- ggplot(dltam_scores) +
  geom_line(aes(x=timestamp, y=score), colour="black") +
  scale_x_date(date_breaks="2 months", date_labels="%b") +
  labs(x="", y="DLTAM score") +
  theme_bw(base_size = 9)

p1 / p2
```
